---
title: 'Skills Test: Question 10'
author: "Charlie Lonergan"
date: "18/10/2021"
output: pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

It is possible to get an approximate value of pi by randomly adding points to a square with a circle inscribed in it.
As more points are added, the ratio of points inside the circle to the total number of points approximates pi / 4.
Using R, please demonstrate how increasing the number of points leads to better estimates of pi. 
Output the results as a graph showing convergence towards the true value.

## Visual representation of pi estimation

```{r}
plot_pi_est = function(n){
  # randomly generate x,y coordinates uniformly
  x = runif(n, min=-1, max = 1)
  y = runif(n, min=-1, max = 1)
  # empty sets to populate with coordinates
  in_x = c()
  in_y = c()
  out_x = c()
  out_y = c()
  # loop through each point
  for (i in 1:n) {
    # if point lies outside circle, add to "out" set
    if (x[i]^2+y[i]^2 > 1) {
      out_x = c(out_x, x[i])
      out_y = c(out_y, y[i])
    # otherwise add to "in" set
    } else {
      in_x = c(in_x, x[i])
      in_y = c(in_y, y[i])
    }
  }
  
  # estimate pi by formula given
  p = 4*length(in_x)/n
  
  # initialize plot
  plot(c(-1, 1), 
       c(-1, 1), 
       main = paste("Estimating pi with ", n, " points\n", "pi ~ ", round(p,3), sep = ""),
       col = "white",
       xlab = "x", 
       ylab = "y", 
       xlim = c(-1,1),
       ylim = c(-1,1))
  # add points
  points(cbind(in_x,in_y), pch = 20, col = "blue")
  points(cbind(out_x,out_y), pch = 20, col = "red")
  # angles for drawing points around the circle
  theta = seq(0, 2*pi, length = 1000)
  # draw the circle, centre (0,0), radius 1
  lines(x = cos(theta), y = sin(theta), lwd = 2.5)
}
```

```{r}
# test function works
plot_pi_est(1000)
```

## pi estimation function for convergence graph

```{r}
estimate_pi = function(n){
  # randomly generate x,y coordinates uniformly
  x = runif(n, min=-1, max=1)
  y = runif(n, min=-1, max=1)
  # empty set to populate with "in" x coordinates
  in_x = c()
  # loop through each point
  for (i in 1:n) {
    # if point lies inside circle, add to "in" set
    if (x[i]^2+y[i]^2 <= 1) {
      in_x = c(in_x, x[i])
    }
  }
  # estimate pi by formula given
  return(4*length(in_x)/n)
}
```

```{r}
# test function works
estimate_pi(1000)
```

## Demonstrate convergence

```{r}
# demonstrate estimates converge to pi
# range of inputs
N = c(10,100,1000,10000,50000)
# vector of pi estimates
pi_est = round(unlist(lapply(N, estimate_pi)),3)
# vector of differences to true pi
diff = round(abs(pi-pi_est),3)
# display convergence table
cbind(N,pi_est,diff)
```

```{r}
# pi estimation plot over several inputs
for (i in N) {
  plot_pi_est(i)
}
```

```{r}
n = 3000
# generate pi estimates up to n
pi_list = unlist(lapply(1:n, estimate_pi))
# plot estimates against n
plot(1:n, 
     pi_list,
     main = paste("Convergence of pi estimates\n", n, "th pi ~ ", round(pi_list[n],3), sep = ""),
     xlab = "n",
     ylab = "pi estimate",
     pch = 4)
# dashed line represents true pi
abline(h = pi, 
       col = "red", 
       lwd = 2.5, 
       lty = 2)
```
