---
title: 'Skills Test: Question 11'
author: "Charlie Lonergan"
date: "18/10/2021"
output: pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data from “students_data.csv”. This dataset has information about the name of students in a given class, their height, weight and smoking habits.

```{r}
# read data
students = read.csv("data/students_data.csv")
```

## Visualise the relationship of the weight of students as a function of the smoking habits.

```{r}
# bmi more appropriate than direct weight 
# since weight alone does not account for height
students$BMI = students$Weight_Kg/(students$Height_cm/100)^2
```

```{r}
# split data by smoking habits
smokers = students[students$SmokingHabits == "Yes",]
occsmokers = students[students$SmokingHabits == "Occasional",]
nonsmokers = students[students$SmokingHabits == "No",]
```

```{r}
# boxplot of each cohort's weight
boxplot(smokers$Weight_Kg, 
        occsmokers$Weight_Kg,
        nonsmokers$Weight_Kg,
        names=c("Yes","Occasional","No"),
        col = c("red", "orange", "green"),
        main = "Boxplot of Weight by Smoking Habits",
        xlab = "Smoking Habits",
        ylab = "Weight (kg)")
```

There appears to be a reduction in weight as smoking decreases.

```{r}
# boxplot of each cohort's BMI
boxplot(smokers$BMI, 
        occsmokers$BMI,
        nonsmokers$BMI,
        names=c("Yes","Occasional","No"),
        col = c("red", "orange", "green"),
        main = "Boxplot of BMI by Smoking Habits",
        xlab = "Smoking Habits",
        ylab = "BMI (kgm-2)")
```

The regular smokers appear to have an increased BMI also. However, the sample sizes are quite small and statistically insignificant. 

## Compute the mean weight of non-smokers, mean weight of occasional smokers, and mean weight of smokers. Does it seem there is a difference in means?

```{r}
mean(smokers$Weight_Kg)
mean(occsmokers$Weight_Kg)
mean(nonsmokers$Weight_Kg)
```

```{r}
mean(smokers$BMI)
mean(occsmokers$BMI)
mean(nonsmokers$BMI)
```

The regular smokers appear to have increased mean weight and BMI.

## Perform a statistical test to see if if the mean weights depend on the smoking habits.

```{r}
# test mean weight 
smok_occs_w = t.test(smokers$Weight_Kg, occsmokers$Weight_Kg, alternative = "g")
smok_nons_w = t.test(smokers$Weight_Kg, nonsmokers$Weight_Kg, alternative = "g")
smok_occs_w
cat("\n")
smok_nons_w
```

```{r}
# check p-values
smok_occs_w$p.value
smok_nons_w$p.value
```

Mean weights are not significantly different. Smoker's mean weight differs to that of the occasional smokers with a p-value of `0.14`, and to that of the non-smokers with a p-value of `0.06`.

Both p-values are greater than `0.05`, so the null hypothesis that the means differ cannot be rejected at `95%` confidence. 

```{r}
# test mean BMI
smok_occs_b = t.test(smokers$BMI, occsmokers$BMI, alternative = "g")
smok_nons_b = t.test(smokers$BMI, nonsmokers$BMI, alternative = "g")
smok_occs_b
cat("\n")
smok_nons_b
```

```{r}
# check p-values
smok_occs_b$p.value
smok_nons_b$p.value
```

Similarly, mean BMI scores do not differ significantly, since associated p-values are `0.12`,`0.11 > 0.05`.

## Test if there is an interaction between weights, height and smoking habits.

```{r}
# boxplot of each cohort's height
boxplot(smokers$Height_cm, 
        occsmokers$Height_cm, 
        nonsmokers$Height_cm,
        names=c("Yes","Occasional","No"),
        col = c("red", "orange", "green"),
        main = "Boxplot of Height by Smoking Habits",
        xlab = "Smoking Habits",
        ylab = "Height (cm)")
```

```{r}
# boxplot of each cohort's weight
boxplot(smokers$Weight_Kg, 
        occsmokers$Weight_Kg,
        nonsmokers$Weight_Kg,
        names=c("Yes","Occasional","No"),
        col = c("red", "orange", "green"),
        main = "Boxplot of Weight by Smoking Habits",
        xlab = "Smoking Habits",
        ylab = "Weight (kg)")
```

```{r}
# boxplot of each cohort's BMI
boxplot(smokers$BMI, 
        occsmokers$BMI,
        nonsmokers$BMI,
        names=c("Yes","Occasional","No"),
        col = c("red", "orange", "green"),
        main = "Boxplot of BMI by Smoking Habits",
        xlab = "Smoking Habits",
        ylab = "BMI (kgm-2)")
```

While it appears that weight increases with increased smoking habits, the students who smoke regularly happen to also be taller than occasional and non-smokers.
Therefore, the smokers also naturally weigh more.
To account for this, we can get a clearer picture of the association of weight and smoking status by measuring the students' BMI, which also takes account of height.

As we have seen, mean BMI does not differ significantly between cohorts, though there appears to be a noticeable subjective increase in BMI with smoking frequency.
In summary, more data is needed to investigate the relationship between weight and smoking habits, since there are only 3 regular and 3 occasional smokers in the study, so these results are not statistically significant.
