---
title: 'Skills Test: Question 12'
author: "Charlie Lonergan"
date: "18/10/2021"
output: pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Breast Cancer Wisconsin (Diagnostic) Data Set can be downloaded from [kaggle.com]{https://www.kaggle.com/uciml/breast-cancer-wisconsin-data}. 
We would like you to perform a 70:30 training/test split, train an XGboost model on the training split, and use it to predict the test data.

To get a good model it may be necessary to perform feature engineering and feature reduction.

We would like any data preprocessing to be done in R, but the model itself should be trained using Python. 
Finally, we would like the results to passed back into R for visualisation. 
Objects can be conveniently passed from R to Python and back using R’s Reticulate package, which interfaces well with RStudio.
Please aim for a high AUC accuracy, ideally over 90%.

```{r, message=FALSE}
# install and load pROC for AUC
#install.packages("pROC")
library(pROC)
```

## Data Preprocessing

```{r}
# read data
wisconsin = read.csv("data/wisconsin_breast_cancer.csv")
```

```{r}
# check data
summary(wisconsin)
```

```{r}
# not sure where "X" column came from
# remove "ID" column and "X" column
wisconsin <- wisconsin[,c(-1,-33)]
summary(wisconsin)
```

```{r}
# check distribution of benign vs malignant
barplot(c(nrow(wisconsin[wisconsin$diagnosis=="M",]),
          nrow(wisconsin[wisconsin$diagnosis=="B",])),
        main = "Barplot of Malignant vs. Benign",
        col = c("blue", "orange"),
        names.arg = c("M","B"),
        xlab = "Diagnosis",
        ylab = "Count")
```

Data slightly imbalanced - may need balancing.

```{r}
# convert M to 1 and B to 0 for XGBoost
wisconsin$diagnosis[wisconsin$diagnosis=="M"] <- 1
wisconsin$diagnosis[wisconsin$diagnosis=="B"] <- 0
# set to numeric
wisconsin$diagnosis <- as.numeric(wisconsin$diagnosis)
```

```{r}
# min/max normalisation function
min_max = function(x){
  return( (x-min(x)) / (max(x)-min(x)) )
}
```

```{r}
# normalise numeric variables
for (i in 2:ncol(wisconsin)) {
  wisconsin[,i] <- min_max(wisconsin[,i])
}
summary(wisconsin)
```

Numeric variables scaled from 0 to 1

```{r}
# split data into training / test sets
set.seed(101)                                           # make partition reproducible
n = floor(0.7*nrow(wisconsin))                          # set training size to 70%
train_rows = sample(seq_len(nrow(wisconsin)), size = n) # set random training samples
training = wisconsin[train_rows,]                       # training partition
testing = wisconsin[-train_rows,]                       # testing partition
```

```{r}
# check "randomness"
mean(training$diagnosis)
mean(testing$diagnosis)
```

```{r}
# export train/test data
write.csv(x = training, file = "./data/wisconsin_training.csv")
write.csv(x = testing, file = "./data/wisconsin_testing.csv")
```

Now run XGBoost by stepping through `Q12_XGBoost.ipynb`.

## Results

```{r}
# read XGBoost output
predictions = read.csv("./data/wisconsin_predictions.csv")
predictions <- predictions[,2]
predictions
```

```{r}
# append to testing set
testing$predictions = predictions
# measure naive accuracy
acc = 100*nrow(testing[testing$diagnosis == testing$predictions,])/nrow(testing)
print(paste("XGBoost prediction accuracy: ", round(acc,2), "%", sep = ""))
```

```{r}
# measure AUC
res = roc(testing$diagnosis,testing$predictions)
auc_acc = auc(res)
auc_acc
```

```{r}
# Visualise AUC
plot.roc(res, main = paste("AUC: ", 100*round(auc_acc,4), "%", sep = ""))
```
