---
title: 'Skills Test: Question 3'
author: "Charlie Lonergan"
date: "18/10/2021"
output: pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Please annotate the code and fill in the blanks

```{r}
#...(...(dplyr))
#...(...(tidyr))
# not sure what double brackets (( )) are for

library(dplyr)
library(tidyr)
```

```{r}
# returns the cumulative distribution function of the hypergeometric distribution
hype<-function(x, k, m, n){
  hyp<-sum(stats::dhyper(x:k, m, n, k))
  return (hyp)
}

# x:k = number of white balls drawn without replacement
# k = number of white balls in urn
# m = number of black balls in urn
# n = number of balls drawn in (0,1,...,k+m)

hype(2,4,4,5)
```

```{r}
d <- readRDS("../d.rds") # load Rdata
head(d)                  # appears to be lists of transcription factors
```

```{r}
dUnnest <- d %>%
  dplyr::mutate(TF = ...(as.character(TF), ","))%>%
  tidyr::...(TF)%>%
  group_by(id) %>%
  dplyr::...(countTotal=n()) 
head(dUnnest)
```

```{r}
dTable = as.data.frame(...(dUnnest$TF))
head(dTable)
```

```{r}
mergedDF<- dplyr::full_join(dUnnest, dTable, by=...)
mergedDF <- mergedDF[,c("TF","countTotal","Freq")]
mergedDF <- mergedDF%>%
  ...(TF)%>%
  mutate(countTotal = max(countTotal))
mergedDF <- mergedDF[!...(mergedDF),]
head(mergedDF)
```

```{r}
colnames(mergedDF)<-c("GeneLocus", "countTotal", "countSig")
mergedDF <- mergedDF %>%
  ...::column_to_rownames(var="GeneLocus")

mVal <-sum(mergedDF$countSig)
nVal<-(sum(mergedDF$countTotal)-mVal)

res1<-unlist(Map(..., mergedDF$countSig, mergedDF$countTotal, ..., nVal))
mergedDF$HyperG_Stats<-res1
mergedDF$FDR_HyperG<-p.adjust(res1,method="BH", n=length(res1))
head(mergedDF)

```

Expected Output
##       id
## 1 id_1_a
## 2 id_2_b
## 3 id_2_b
## 5 id_3_b
## 6 id_3_b
## 4 id_4_b
##                                                                                                                                                                                                                                                                                                                                                                                                 TF
## 1                                                                                                                                                                                                                                                                                                                                            CHD8,ESR1,RELA,MAZ,XBP1,ZNF19,CTCF,STAG1,ZFX,WT1,CTCF
## 2                                                                                                                                                                                                                                                                                                                                                                               HOXB13,CEBPG,CEBPB
## 3                                                                                                                                                                                                                                       ZNF362,ZNF384,OGG1,BCL6,ZBTB40,ELF4,C11orf30,RXR,MAFG,AR,ZNF316,MAFK,GRHL3,DUX4,TEAD1,PRDM1,ZNF512,ZNF366,FOXA1,RELA,ZNF143,XBP1,GTF2B,TEAD3,CREB1,MYC,ZFX
## 5                                                                                                                                                                                                                                                                                                                                                                              SP140,ZBTB33,ZNF384
## 6 ZNF121,BCOR,ZNF770,SP140,BRD4,WT1,CEBPB,CEBPG,CEBPA,NFIL3,LMO2,HLF,MYC,FLI1,CBX2,ELF3,BRD4,CHD7,SOX2,SCRT2,TAL1,RELA,JUN,PAF1,FOXO1,RBPJ,JUNB,FOXP2,NR3C1,RUNX1,PAF1,ZNF341,PGR,ZFX,BRD2,MAX,CDK8,MYC,BRD4,ZBTB48,AR,PRPF4,NIPBL,WT1,RELA,ERG3,CHD8,SMARCA4,ZFHX2,SP140,ERG2,SRF,NEUROG2,RAD21,STAG1,ZNF592,CTCF,KLF9,XBP1,MYCN,RUNX3,YY1,PHIP,SOX4,TP53,EBF1,SMC1A,RUNX2,POU5F1,GRHL3,CTCFL,AHR
## 4                                                                                                                                                                                                                                                                                                                                                                                      GTF2B,FOXA1
## # A tibble: 6 × 3
## # Groups:   id [1]
##   id     TF    countTotal
##   <chr>  <chr>      <int>
## 1 id_1_a CHD8          11
## 2 id_1_a ESR1          11
## 3 id_1_a RELA          11
## 4 id_1_a MAZ           11
## 5 id_1_a XBP1          11
## 6 id_1_a ZNF19         11
##    Var1 Freq
## 1   AHR    4
## 2    AR    4
## 3  ARNT    1
## 4 BACH1    1
## 5  BCL6    4
## 6 BCL6B    1
## # A tibble: 6 × 3
## # Groups:   TF [6]
##   TF    countTotal  Freq
##   <chr>      <int> <int>
## 1 CHD8         102     4
## 2 ESR1         102     2
## 3 RELA         102     7
## 4 MAZ           11     1
## 5 XBP1         102     4
## 6 ZNF19         11     1
##       countTotal countSig HyperG_Stats FDR_HyperG
## CHD8         102        4   0.16912031  0.8876704
## ESR1         102        2   0.63832844  0.8876704
## RELA         102        7   0.00579439  0.4114017
## MAZ           11        1   0.20931099  0.8876704
## XBP1         102        4   0.16912031  0.8876704
## ZNF19         11        1   0.20931099  0.8876704





















