---
title: 'Skills Test: Question 7'
author: "Charlie Lonergan"
date: "18/10/2021"
output: pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lions are born with pink noses. As they grow, black spots start to appear in their nose. 
Data on this can be pulled using the following R code.

```{r}
lion <- read.csv(url("http://www.zoology.ubc.ca/~schluter/WhitlockSchluter/wp-content/data/chapter17/chap17e1LionNoses.csv"))
```

After pulling the data ‘lion’, plot the relationship between age and proportion of black spots. 
Next, predict how old a lion is that has a 0.5 proportion of black in its nose.

```{r}
# plot data
plot(x = lion$ageInYears, 
     y = lion$proportionBlack,
     main = "Proportion of Black Spots vs. Age",
     col = "red",
     pch = 18,
     xlim = c(0,14),
     ylim = c(0,0.8),
     xlab = "Age (Years)", 
     ylab = "Black spot proportion")
```

Scatterplot appears to describe a linear relationship, though it is possible a log-linear model is more appropriate.

```{r}
# predicting age as a function of proportion
print("ageInYears ~ proportionBlack")
summary(lm(lion$ageInYears ~ lion$proportionBlack))$coefficients
cat("\n")
print("ageInYears ~ log(proportionBlack)")
summary(lm(lion$ageInYears ~ log(lion$proportionBlack)))$coefficients
cat("\n")
print("log(ageInYears) ~ proportionBlack")
summary(lm(log(lion$ageInYears) ~ lion$proportionBlack))$coefficients
```

"ageInYears ~ log(proportionBlack)" has lowest associated p-values for both gradient and intercept, so will choose this transformation as the basis of the log-linear model.

```{r}
# log transform black spot proportions
lion$logProportion = log(lion$proportionBlack)
# build linear model: log proportion as a predictor of age
lion_fit = lm(lion$ageInYears ~ lion$logProportion)
summary(lion_fit)
```

Predict age of lion with 50% black spots.

```{r}
lp = log(0.5)                  # log transform 
c = lion_fit$coefficients[[1]] # intercept
m = lion_fit$coefficients[[2]] # gradient

# interpolate how old a lion is that has a 0.5 proportion of black in its nose.
middle_age = m*lp + c
middle_age
```

Visualise interpolation.

```{r, warning=FALSE}
# scatter plot
plot(x = lion$ageInYears,
     y = lion$logProportion,
     main = "Predicting lion age with 50% black spots",
     col = "red",
     pch = 18,
     xlab = "Age (Years)", 
     ylab = "ln( Black spot proportion )", 
     xlim = c(0,14),
     ylim = c(-2.5,0))#
# indicate interpolation at 50% black spots
abline(v = middle_age, h = lp, lty = 3)
points(middle_age, lp)
# max/min values for interpolation
y = range(lion$logProportion)
# predict() returns the min/max age values predicted by lm()
x = range(predict(lion_fit, data.frame(logProportion = y)))
# lines() draws a line between the min/max x,y values
lines(x = x, y = y)
```

