---
title: 'Skills Test: Question 9'
author: "Charlie Lonergan"
date: "18/10/2021"
output: pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You are sitting in a room and there is a party going on in the room next door. 
Through the walls you hear everyone clink glasses (everyone clinks glasses with everyone else once). 
You hear 28 distinct clinks. How many people were at the party? 
Write a function to calculate how many clinks would be generated for any number of guests. 

## Create functions

This problem is an example of a [complete graph](https://en.wikipedia.org/wiki/Complete_graph). 
The number of people can be represented by nodes, and the number of clinks by edges.

For `n` nodes in a complete graph, the number of edges is `n(n-1)/2`, so given `k` edges there are `1+sqrt(1+8k)/2` edges (by the quadratic formula).

```{r}
# function returns how many clinks generated for any number of guests. 
num_edges = function(n){
  return(n*(n-1)/2)
}
```

```{r}
# inverse function returns how many guests generated number of clinks heard.
num_nodes = function(k){
  return((1+sqrt(1+8*k))/2)
}
```

```{r}
print(paste("There were ", num_nodes(28), " people at the party", sep = ""))
```

```{r}
# calculate number of clinks for any number of people /
# number of people for any number of clinks
k = num_edges(1:25)
n = num_nodes(k)
cbind(n,k)
```

